---
title: "Homework 5 Solution"
author: "Safial Islam Ayon"
date: "November 28, 2021"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Probability distributions

Explore the functions *rnorm*, *rt*, *runif*, *rbinom* in R that allow you to generate random numbers from the normal, t-, uniform, and binomial distribution. Compute them with different values, and inspect histograms to visualize their distribution.

```{r}
#Normal Distribution

ND1 <- rnorm(n = 500, mean = 1, sd = 0.5)
ND2 <- rnorm(n = 500, mean = 2, sd = 1.0)

#Visualize the distribution

par(mfrow = c(1, 2))
hist(ND1, freq = FALSE, xlim = c(-2, 4), ylim = c(0, 1))
hist(ND2, freq = FALSE, xlim = c(-3, 6), ylim = c(0, 1))


```
```{r}

#t-distributions

TD1 <- rt(n = 1000, df = 600)
TD2 <- rt(n = 1000, df = 400)
TD3 <- rt(n = 1000, df = 200)

#Visualize the distribution

par(mfrow = c(1, 3))
hist(TD1, freq = FALSE, xlim = c(-5, 5), ylim = c(0, 0.5))
hist(TD2, freq = FALSE, xlim = c(-5, 5), ylim = c(0, 0.5))
hist(TD3, freq = FALSE, xlim = c(-5, 5), ylim = c(0, 0.5))

```

```{r}

#Uniform Distributions

UD1 <- runif(n = 1000, min = 0, max = 5)
UD2 <- runif(n = 1000, min = -2, max = 2)
UD3 <- runif(n = 1000, min = -5, max = -1)

#Visualize the distribution
par(mfrow = c(1, 3))
hist(UD1, freq = FALSE, xlim = c(-2, 6), ylim = c(0, 0.4))
hist(UD2, freq = FALSE, xlim = c(-5, 5), ylim = c(0, 0.4))
hist(UD3, freq = FALSE, xlim = c(-6, 0), ylim = c(0, 0.4))

```


```{r}

#Normal distributions

ND1 <- rnorm(n = 5000, mean = 0.5, sd = 1.5)
ND2 <- rnorm(n = 5000, mean = 0.5, sd = 1.5)
ND3 <- rnorm(n = 5000, mean = 0.5, sd = 1.5)

#visualize it, plot histograms with the relative frequency on the Y axis
par(mfrow = c(1, 3))
hist(ND1, freq = FALSE, xlim = c(-10, 10), ylim = c(0, 0.4))
hist(ND2, freq = FALSE, xlim = c(-10, 10), ylim = c(0, 0.4))
hist(ND3, freq = FALSE, xlim = c(-10, 10), ylim = c(0, 0.4))

```

## Exercise 2: Odds ratio

In the KiGGS dataset:

a) Compute the proportion of mothers that had hypertension during pregnancy. Use the variable 'e0155' which has values "Ja" (yes), "Nein" (No) and "Weiß nicht" (don't know).
b) Create a new variable that is 0 or 1 depending on whether the children are small or tall (think of a good way how to do this) based on the variable 'groeB' (body height). 
c) Then compute the odds ratio that the mother had hypertension during pregnancy (e0155 == "Ja" (yes), versus e0155 == "Nein" (no)) of tall vs. small children.

```{r}
load("KiGGS03_06.RData")

Kiggs <- KiGGS03_06

table(Kiggs$e0155)

#Proportion of mothers that had hypertension during pregnancy = (202)/(202+2305+37) = 0.0794 or 7.9% 

```

```{r}
str(Kiggs$groeB)
typeof(Kiggs$groeB)
Kiggs$groeB <- as.numeric(as.character(Kiggs$groeB))

tapply(Kiggs$groeB, Kiggs$age2, mean, na.rm = TRUE)

Kiggs$tall <- NULL

for(i in 1:length(levels(Kiggs$age2)))
{
  n <- (Kiggs$age2 == levels(Kiggs$age2)[i])
  Kiggs$tall[n] <- ifelse(Kiggs$groeB[n] > mean(Kiggs$groeB[n], na.rm = TRUE), 1, 0)
}

Kiggs$tall <- factor(Kiggs$tall, labels = c("short", "tall"))
table(Kiggs$tall)
                        
```


```{r}
table(Kiggs$e0155)
Kiggs$e0155[Kiggs$e0155 == "Weiß nicht"] <- NA
Kiggs$e0155 <- droplevels(Kiggs$e0155)

table(Kiggs$e0155, Kiggs$tall)

tall_prob <- 83/(83+857)
short_prob <- 118/(118+1404)

tall_odds <- tall_prob/(1-tall_prob)
tall_odds
short_odds <- short_prob/(1-short_prob)
short_odds

tall_odds/short_odds

```



## Exercise 3 (optional): Confidence intervals

Look at the hypertension variable from exercise 2. Use the binom::binom.confint and the questionr::odds.ratio functions to compute the estimates of the proportion and odds ratio as well as their confidence intervals. (you need to download and load these packages at first).

```{r}
library("binom")
library("questionr")

binom::binom.confint(83, 83+857)
binom::binom.confint(118, 118+1404)

questionr::odds.ratio(table(Kiggs$e0155, Kiggs$tall))
```

## Exercise 4 (optional, advanced): Bootstrap

Adapt the bootstrap implementation in R_5b_estimation_bootstrap.Rmd to compute the bootstrap estimate of the standard error of the variance of a normally-distributed and a t-distributed variable. Are they similar?

```{r}

set.seed(1)
n <- 100
sigma <- 1
X <- rnorm(n = n, mean = 0, sd = sigma)

mean_BS <- NULL

k = 1000
for(i in 1:k)
{
  X_sample_i <- sample(X, size = n, replace = TRUE)
  mean_BS[i] <- mean(X_sample_i)
}

SE_BS_mean <- sd(mean_BS)
SE_BS_mean

Y <- rt(n = n, df = 500)

mean_t <- NULL
k = 1000
for(i in 1:k)
{
  y_sample_i <- sample(Y, size = n, replace = TRUE)
  mean_t[i] <- mean(y_sample_i)
}

SE_t_mean <- sd(mean_t)
SE_t_mean

difference = SE_t_mean - SE_BS_mean
difference

#They are not similar.
```