---
title: "Homework 7"
author: "Stefan Konigorski"
date: "December 9, 2021"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Preparations

Load KiGGS data:

```{r}
dat_link <- url("https://www.dropbox.com/s/pd0z829pv2otzqt/KiGGS03_06.RData?dl=1")
load(dat_link)
kiggs <- KiGGS03_06
```

## Exercise 1: Multiple imputation

a) In the KiGGS dataset, choose 2 metric variables of your choice.
b) Compute the correlation between these two variables
c) Perform a multiple imputation using the mice function with the 'pmm' method and otherwise default settings.
d) Compute the correlation again on the multiple imputed datasets.


```{r}

var1 = kiggs$wKGSLab
var2 = kiggs$wKGSVitD

```

```{r}

#Pearson correlation test
pc <- cor.test(var1, var2, method = "pearson")
pc

#In the result
#t is the t-test statistic value;
#df is the degrees of freedom;
#p-value is the significance level of the t-test;
#conf.int is the confidence interval of the correlation coefficient at 95%;
#sample estimates is the correlation coefficient.


#Kendall rank correlation test
krc <- cor.test(var1, var2, method="kendall")
krc

#In the result, tau is the Kendall correlation coefficient.

#Spearman rank correlation coefficient
src <- cor.test(var1, var2, method = "spearman")
src

#In the result, rho is the Spearman’s correlation coefficient.

```


```{r}

library(mice)

#First I try the mice function of the whole kiggs dataset. But it doesnot work. It shows the following errors: "Error: cannot allocate vector of size 4.5 Gb" 
#Then I create a new dataframe using the variable I take at first. 

var = data.frame(var1, var2)

imp1 <- mice(var, m = 5) #The default method of imputation in the MICE package is PMM and the default number of imputations is 5
imp1

```


```{r}

temp1 = imp1$imp$var1
temp2 = imp1$imp$var2

temp3 <- temp2[-c(3255:7625), ] #Reshape the temp2 dataframe because temp1 and temp2 dataframes dimensions are not same. 

correlation <- cor(temp1,temp3) #Correlation between two variables
correlation

```