---
title: "Homework 12"
author: "Safial Islam Ayon"
date: "February 3, 2022"
output:
  html_document: default
---

```{r}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Linear mixed model for hierarchical data

Load the HRS dataset and do the following preparations:

```{r}
library(readr)
dat_link <- url("https://www.dropbox.com/s/z281ijk81ohyg9v/hrs.RData?dl=1")
load(dat_link)

# Preparations
# rename variable at first timepoint
names(hrs)[5] <- "R1IMRC"
# define variable at second timepoint - Integrate the variables R2AIMRC10 and R2HIMRC20, which both capture different subgroups of the sample at the second timepoint.
table(is.na(hrs$R2AIMRC10)) # -> only ~7000 of ~36000 observations
table(is.na(hrs$R2HIMRC20)) # -> only ~10000 of ~36000 observations
table(!is.na(hrs$R2HIMRC20) & !is.na(hrs$R2AIMRC10)) # -> no person has a value at both variables, i.e. the variables complement each other
hrs$R2IMRC <- rowSums( cbind (hrs$R2AIMRC10, hrs$R2HIMRC20), na.rm=TRUE)
hrs$R2IMRC[!(!is.na(hrs$R2HIMRC20) | !is.na(hrs$R2AIMRC10))] <- NA
table(is.na(hrs$R2IMRC))
# Rescale R1IMRC and R2IMRC onto the same scale as all other variables (measures 3-12 were on a scale of 1-10, measures 1-2 on the scale 1-20)
hrs$R1IMRC <- hrs$R1IMRC/2
hrs$R2IMRC <- hrs$R2IMRC/2

# Extract variables for first investigation
data_IMRC <- data.frame(HHID = hrs$HHID, PN = hrs$PN, HHID_PN = paste(hrs$HHID, hrs$PN, sep = "_"),
                        T1 = hrs$R1IMRC, T2 = hrs$R2IMRC, T3 = hrs$R3IMRC, T4 = hrs$R4IMRC, T5 = hrs$R5IMRC, T6 = hrs$R6IMRC,
                        T7 = hrs$R7IMRC, T8 = hrs$R8IMRC, T9 = hrs$R9IMRC, T10 = hrs$R10IMRC, T11 = hrs$R11IMRC, T12 = hrs$R12IMRC)
```

Answer the following questions: 

a) How many observations are there in the dataset?

```{r}
dim(data_IMRC)

nrow(data_IMRC)

## Result 36167    15
## So 36167 observations and 15 variables in the dataset.
```

b) How many households are there in the dataset (observations are clustered in households)?

```{r}
## We can do this two ways

## First Way
## Make data frame using the "households", then calculate the observation.
house_count <- as.data.frame(table(data_IMRC$HHID))
dim(house_count)

## Second way
## Using the n_distinct function
library(dplyr, warn.conflicts = FALSE)
n_distinct(data_IMRC$HHID)

## Both the cases the result is 22996
## That means there are 22996 households clustered among the the 36167 total data.  
```

c) Now, compute a linear regression model to predict T2 by T1, i.e. the cognitive performance at the second timepoint by the cognitive performance at the first timepoint. Also, compute two linear mixed models, one with random intercept, one with random intercept+slope to predict T2 by T1. Compare the results, i.e. the estimated effect of T1 on T2. Are they similar, or are they very different?


```{r}

## linear regression model
model1 <- lm(T2 ~ T1, data = data_IMRC)
summary(model1)

## The coefficient Estimate contains two rows; the first one is the intercept. 
## It takes an average of cognitive performance in the dataset 1.94935 between two timepoint.
## The second one is slpoe. 
## Here the cognitive performance goes up 0.53446.

```

```{r}
## linear regression model with random intercept
library(lme4)
model2 = lmer(T2 ~ T1 + (1|HHID), REML = T, data = data_IMRC)
summary(model2)

## If we look at the summary output we see under the Random Effects that the residual variance on the class level 0.2228 and residual variance on the first level is 1.7701. This means that the intraclass correlation (ICC) is 0.2228/(1.7701+0.2228)=.1118.
## Under Fixed Effects the estimate of the intercept is reported, which is 1.96389.
```
```{r}
## linear regression model with random intercept+slope 
library(nlme)
model3 = lmer(T2 ~ T1 + (1 + T1|PN), data = data_IMRC, REML = T)
summary(model3)

## The variance of the intercept and for the slope of T1 is 0.03755 and 0.00163 respectively.

## Random intercept or random slope model?
## A random intercept model is more parsimonious. Random slope model is more flexible as it allows the effect of days of sleep deprivation to vary across participants.

m1 <- model2
m2 <- model3

anova(m1, m2)
## Results from the likelihood ratio test shows that the random slope model fits the data significantly better than the random intercept model. 
```

```{r}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```