---
title: "Homework 4"
author: "Stefan Konigorski"
date: "November 18, 2021"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Descriptive statistics - example 1

Create a summary table with descriptive statistics of the esoph dataframe in the datasets package. For example by using the summary_table function in the qwraps2 package.

Alternatively, only compute the descriptive statistics of interest, and put them into a table in Word (or Excel, or another program).

```{r}
# Load the data
library(datasets)
?esoph
dat1 <- esoph
```

## Exercise 2: Descriptive statistics - example 2 (optional, more advanced)

Create a summary table with descriptive statistics of the UKLungDeaths data in the datasets package. For example by using the summary_table function in the qwraps2 package. 

Alternatively, only compute the descriptive statistics of interest, and put them into a table in Word (or Excel, or another program).

```{r}
# Load the data
library(datasets)
?UKLungDeaths 
```

Hint: the data consists of three datasets.

## Exercise 3: Sample size calculation (optional - only solvable if you have taken some statistics course in the past)

Aim: Compute the sample size using R or G*Power.

Study question: Investigate if biking to the HPI helps you to better concentrate in class or not. 

Think about which study design you could use for this study, and describe it.

Think about which statistical test you could use for investigating this research question, and describe it.

Do a sample size calculation and answer the question: which sample size is necessary, to find a true effect with 80\% power at $\alpha=0.05$?