---
title: "Homework 4"
author: "Safial Islam Ayon"
date: "November 21, 2021"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Descriptive statistics - example 1

Create a summary table with descriptive statistics of the esoph dataframe in the datasets package. For example by using the summary_table function in the qwraps2 package.

Alternatively, only compute the descriptive statistics of interest, and put them into a table in Word (or Excel, or another program).

```{r}
# Load the data
library(datasets)
?esoph
dat1 <- esoph
#dat1

#Solution
library(qwraps2)
options(qwraps2_markup = "markdown")

library(dplyr)

dat1 %>%
  dplyr::select(.data$tobgp, .data$agegp, .data$alcgp, .data$ncases, .data$ncontrols) %>%
  dplyr::group_by(agegp) %>%
  summary_table(.) %>%
  print(rtitle = "Summary Descriptive Statistics Table", cnames = c("Age: 25-34    ", "Age: 35-44    ", "Age: 45-54    ", "Age: 55-64    ", "Age: 65-74    ", "Age: 75+    "))

```


## Exercise 2: Descriptive statistics - example 2 (optional, more advanced)

Create a summary table with descriptive statistics of the UKLungDeaths data in the datasets package. For example by using the summary_table function in the qwraps2 package. 

Alternatively, only compute the descriptive statistics of interest, and put them into a table in Word (or Excel, or another program).

```{r}
# Load the data
library(datasets)
?UKLungDeaths 

#Solution
library(qwraps2)

ldeaths
fdeaths
mdeaths

```

Hint: the data consists of three datasets.

## Exercise 3: Sample size calculation (optional - only solvable if you have taken some statistics course in the past)

Aim: Compute the sample size using R or G*Power.

Study question: Investigate if biking to the HPI helps you to better concentrate in class or not. 

Think about which study design you could use for this study, and describe it.

Think about which statistical test you could use for investigating this research question, and describe it.

Do a sample size calculation and answer the question: which sample size is necessary, to find a true effect with 80\% power at $\alpha=0.05$?

solution:
I will use "Analytic studies - N-of-1 trial" study design, because, the goal of an n-of-1 trial is to determine the optimal or best intervention for an individual person using objective data-driven criteria [1].This type of study has enabled practitioners to achieve experimental progress without the overwhelming work of designing a group comparison study. It can be very effective in confirming causality [2].

[1] https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3118090/
[2] https://en.wikipedia.org/wiki/N_of_1_trial


Absolute frequency could use for investigating the research question, because Absolute frequency is a statistical term describing the number of times a particular piece of data or a particular value appears during a trial or set of trials. Essentially, absolute frequency is a simple count of the number of times a value is observed. Absolute frequency is usually expressed as a whole number and is considered a very basic level of statistical analysis [3].

[3] https://www.investopedia.com/terms/a/absolute_frequency.asp


In G*power, first select statistical test = Linear multiple regression: Fixed model, single regression coefficient, Tail(s) = One, Effect size f^2 = 0.15, alpha = 0.05, Power = 0.8, and number of predictors = 2, then click calculate. Then found that total sample size = 43.   
