---
title: "Homework 13"
author: "Safial Islam Ayon"
date: "February 10, 2022"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Survival analysis

Choose one other provided dataset in the survival package (or any other suitable dataset) which contains time-to-event data. Use the survfit function to compute the Kaplan-Meier estimate of the survival function and plot it. Then choose a variable with 2 or more categories in the dataset and perform a log-rank test. Interpret the results: is there are difference in survival between the groups?

```{r}
library(survival)

s <- Surv(lung$time, lung$status)
survfit(s~1)
survfit(Surv(time, status)~1, data=data_new)
sfit <- survfit(Surv(time, status)~1, data=data_new)
sfit
```

```{r}
#Fit a survival curve to the data it’s pretty easy to visualize it with a Kaplan-Meier plot
sfit <- survfit(Surv(time, status)~sex, data=data_new)
plot(sfit)
```

```{r}
#Add confidence intervals, show the p-value for the log-rank test, show a risk table below the plot, and change the colors and the group labels.
library(survminer)
ggsurvplot(sfit, conf.int=TRUE, pval=TRUE, risk.table=TRUE,legend.labs=c("Male", "Female"), legend.title="Sex", palette=c("dodgerblue2", "orchid2"), title="Kaplan-Meier Curve for Lung Cancer Survival", risk.table.height=.15)

```

```{r}
surv_diff <- survdiff(Surv(time, status) ~ sex, data = data_new)
surv_diff

#With one degree of freedom, the Chi-Squared test statistic is 10.3 and the associated p-value is 0.001. We reject the null hypothesis since the p-value is less than 0.05.

#In other words, we have enough data to conclude that the two groups have statistically significant differences in survival.
```